```{r setup, include=FALSE}
library(scales)
library(tidyverse)
library(ggplot2)
library(shiny)
library("plotly")
library(lubridate)
require(rgl)
require(akima)
library(dplyr)
library(DT)
library(readxl)
library(neuralnet)
library(GGally)
library(janitor)
df <- read_excel("MFF_Singular_Sheet.xlsx")
col_names = c("Institution", "Degree Level", "Major", "Major code", "Cohort Year", "Graduates (with full-time wages)", "25th percentile earnings: full-time workers", "Median earnings: full-time workers", 
"75th percentile earnings: full-time workers", "Percent of graduates continuing education", "Percent working full-time", "Median earnings: Bachelor's degree earners with no additional credential", "Percent of bachelor's degree earners completing: No additional credential", "Median earnings: Bachelor's degree earners with additional master's degree", "Percent of bachelor's degree earners completing: Additional masters degree", "Median earnings: Bachelor's degree earners with additional doctorate degree", "Percent of bachelor's degree earners completing: Additional doctorate degree", "Formatted Major Code")
crosswalk <- read_excel("CIP_SOC.xlsx")
col_names = c("SOC2018Code", "CIP2020Code")
jobs <- read_excel("Florida Jobs.xlsx")
col_names = c("AREA", "AREA_TITLE", "AREA_TYPE", "PRIM_STATE", "NAICS", "NAICS_TITLE", "I_GROUP", "OWN_CODE", "OCC_CODE", "OCC_TITLE", "O_GROUP", "TOT_EMP", "EMP_PRSE", "JOBS_1000", "LOC_QUOTIENT", "PCT_TOTAL", "PCT_RPT", "H_MEAN", "A_MEAN", "MEAN_PRSE", "H_PCT10", "H_PCT25", "H_MEDIAN", "H_PCT75", "H_PCT90", "A_PCT10", "A_PCT25", "A_MEDIAN", "A_PCT75", "A_PCT90", "ANNUAL", "HOURLY")
group_codes <- read_excel("Major_Group_Codes.xlsx")
col_names = c("Major group code", "Major group")

df1 <- df
df1$"Major group code" = substr(df1$"Major code",1,3)
df1
merged <- left_join(x = df1, y = group_codes, by = "Major group code")
df<-filter(df, df$"Median earnings: full-time workers" != "**")
merged<-filter(merged, merged$"Median earnings: full-time workers" != "**")
jobs<-filter(jobs, A_MEDIAN != "*")
jobs<-filter(jobs, A_MEDIAN != "#")
filtered_data <-  df %>% filter(df$"Degree Level" == "Bachelors",
                           df$"Cohort Year" == "1 year after graduation",
                           df$"Percent of graduates continuing education" != '**')
```

> This search is a simplified version of the final product. These results will produce the averaged earnings of full-time workers from a selection of Florida universities. In addition, occupations are matched to prior results through the Standard Occupational Classification (SOC) Code system and their corresponding salaries are provided.

```{r}
salary_search <- function() {

  routes <<- (data.frame(unique(df['Institution'])))
  
  ui <- fluidPage(
    titlePanel("Salary and Occupation Match"),
    sidebarPanel(
      selectInput("university_input", "Select the university", choices = routes),
      selectInput("major_input", "Select the major", c("")),
      selectInput("degree_input", "Select the degree level", c("")),
      selectInput("cohort_input", "Select the cohort year", c(""))
     ),
    mainPanel(width = 12, DT::DTOutput("mytable"), textOutput("text1"), textOutput("text2"), textOutput("text3"), htmlOutput("text4"), tags$head(tags$style("#text1{color: black; font-size: 20px; font-style: bold;}")), tags$head(tags$style("#text2{color: black; font-size: 20px; font-style: bold;}")), tags$head(tags$style("#text3{color: black; font-size: 20px; font-style: bold;}")), tags$head(tags$style("#text4{color: black; font-size: 20px; font-style: bold;}")))
  )
  
  server <- function(input, output, session) {

    observeEvent(input$university_input,
               {
                 w <<- input$university_input
                 updateSelectInput(session, "major_input",
                                   choices =unique(df[df$Institution == input$university_input, ]$Major),
                 )
               }
  )
  observeEvent(input$major_input,
               {
                 x <<- input$major_input
                 updateSelectInput(session, "degree_input",
                                   choices = unique(df[(df$Institution == input$university_input & df$Major == input$major_input), ]$"Degree Level"),
                 )
               }
  )
  
  observeEvent(input$degree_input,
               {
                 y <<- input$degree_input
                 updateSelectInput(session, "cohort_input",
                                   choices = unique(df[(df$Institution == input$university_input & df$Major == input$major_input & df$"Degree Level" == input$degree_input), ]$"Cohort Year"),
                 )
               }
  )
  
  observe(z <<- input$cohort_input)
  
  output$mytable <- DT::renderDT({
     df %>% filter(Institution == input$university_input,
                  Major == input$major_input,
                  df$"Degree Level" == input$degree_input,
                  df$"Cohort Year" == input$cohort_input)})
  
  output$text1 <- renderText({ 
    paste("The 25th percentile for this selection makes $", unique(df[(df$Institution == input$university_input & df$Major == input$major_input & df$"Degree Level" == input$degree_input & df$"Cohort Year" == input$cohort_input), ]$"25th percentile earnings: full-time workers"), sep = "")})
  
    output$text2 <- renderText({ 
    paste("The median for this selection makes $", unique(df[(df$Institution == input$university_input & df$Major == input$major_input & df$"Degree Level" == input$degree_input & df$"Cohort Year" == input$cohort_input), ]$"Median earnings: full-time workers"), sep = "")})
    
  output$text3 <- renderText({ 
    paste("The 75th percentile for this selection makes $", unique(df[(df$Institution == input$university_input & df$Major == input$major_input & df$"Degree Level" == input$degree_input & df$"Cohort Year" == input$cohort_input), ]$"75th percentile earnings: full-time workers"), sep = "")})
  
  output$text4 <- renderText({
    req(input$cohort_input)
    CIP <- filter(crosswalk, CIP2020Code == (unique(df[(df$Institution == input$university_input & df$Major == input$major_input & df$"Degree Level" == input$degree_input & df$"Cohort Year" == input$cohort_input), ]$"Formatted Major Code")))
    selected_job_codes <-  c(CIP$SOC2018Code)
    job_codes <- filter(jobs, is.element(OCC_CODE, selected_job_codes))
    paste0("<p>","The occupation is ", job_codes$OCC_TITLE, " and the median amount is $", job_codes$A_MEDIAN, "<p>")})
  }
  
  shinyApp(ui, server)
}

salary_search()
```

> These results will compare a selected major and their averaged salary and how they fare based on the Florida university attended.

```{r}
compare_search <- function() {

  routes <<- (data.frame(unique(df['Major'])))
  
  ui <- fluidPage(
    titlePanel("Compare Search"),
    sidebarPanel(
      selectInput("major_input", "Select the major", choices = routes),
      selectInput("degree_input", "Select the degree level", c("")),
     ),
    plotOutput("barplot"),
  )
  
  server <- function(input, output, session) {

    observeEvent(input$major_input,
               {
                 major_selection <<- input$major_input
                 updateSelectInput(session, "degree_input",
                                   choices = unique(df[df$Major == input$major_input, ]$"Degree Level"),
                 )
               }
  )
  observe(degree_selection <<- input$degree_input)
  
  output$barplot <-renderPlot({
    req(input$degree_input)
    selection <- df %>% filter(Major == input$major_input, df$"Degree Level" == input$degree_input)
    cohort_order <- c("1 year after graduation", "5 years after graduation", "10 years after graduation")
    selection$"Cohort Year" <- factor(selection$"Cohort Year", levels=cohort_order)
    Institutions <-  c(selection$Institution)
    Cohort_Years <-  c(selection$"Cohort Year")
    #This avoids any sort of scientific notation being printed out when the columns are labeled
    Non_Numeric_Median_Earnings <- c(selection$"Median earnings: full-time workers")
    Median_Earnings <- as.numeric(Non_Numeric_Median_Earnings)
    data <- data.frame(Institutions,Cohort_Years,Median_Earnings)
    
    ggplot(data, aes(fill = Cohort_Years, x = Institutions, y = Median_Earnings)) + geom_bar(position="dodge", stat="identity") + geom_text(
    aes(label = Non_Numeric_Median_Earnings),
    color = "black", size = 4,
    vjust = 1.5, position = position_dodge(.9))
  })
  
  }
  
  shinyApp(ui, server)
}

compare_search()
```

> These results will produce the rate of those with a bachelors degree who continue their education during cohort year 1.

```{r}
continuing_education_search <- function() {

  routes <<- (data.frame(unique(filtered_data['Major'])))
  
  ui <- fluidPage(
    titlePanel("Postgraduate Education"),
    sidebarPanel(
      selectInput("major_input", "Select the major", choices = routes)
     ),
    plotOutput("barplot"),
  )
  
  server <- function(input, output, session) {

  output$barplot <-renderPlot({
    req(input$major_input)
    selection <- filtered_data %>% filter(Major == input$major_input)
    Unordered_Institutions <-  c(selection$Institution)
    Percent_Continuing_Education <-  c(selection$"Percent of graduates continuing education")
    Institutions <- reorder(Unordered_Institutions, -as.numeric(Percent_Continuing_Education))
    
    Percent_Continuing_Education <- as.numeric(Percent_Continuing_Education)
    data <- data.frame(Institutions, Percent_Continuing_Education)
    
    ggplot(data, aes(fill = Institutions, x = Institutions, y = Percent_Continuing_Education)) + geom_bar(position="dodge", stat="identity") + geom_text(
    aes(label = scales::percent(Percent_Continuing_Education)),
    colour = "black", size = 4,
    vjust = 1.5, position = position_dodge(.9)) + scale_y_continuous(labels = scales::percent)
  })
  
  }
  
  shinyApp(ui, server)
}

continuing_education_search()
```

```{r}
df_nn <- merged
df_nn[df_nn == "**"] <- NA
#Any column that has null values has to be omitted
#Other columns that repeat similar information are removed as well
df_nn <- subset(df_nn, select = c("Institution", "Degree Level", "Major", "Major code", "Cohort Year", "Graduates (with full-time wages)", "25th percentile earnings: full-time workers", "Median earnings: full-time workers", 
"75th percentile earnings: full-time workers","Percent working full-time", "Major group"))

#Takes care of the '25 group that is not defined
df_nn$"Major group"[is.na(df_nn$"Major group")] <- 'Library Science'

#Changing certain columns from chr to num or factors
df_nn$"Major code" = substr(df_nn$"Major code",2,7)
i <- c(4, 6, 7, 8, 9, 10) 
df_nn[ , i] <- apply(df_nn[ , i], 2,
                    function(x) as.numeric(as.character(x)))
df_nn <- df_nn %>% mutate_if(is.character, as.factor)

#Prep column names for model.matrix
colnames(df_nn)[2] <- "Degreelevel"
colnames(df_nn)[4] <- "Majorcode"
colnames(df_nn)[5] <- "Cohortyear"
colnames(df_nn)[6] <- "Graduates"
colnames(df_nn)[7] <- "Twentyfifthpercentile"
colnames(df_nn)[8] <- "Median_earnings"
colnames(df_nn)[9] <- "Seventyfifthpercentile"
colnames(df_nn)[10] <- "Percentfulltime"
colnames(df_nn)[11] <- "Majorgroup"

#Setting seed for reproducibility
set.seed(5)

#Create Dummy Variables
m <- model.matrix( 
  ~ Median_earnings + 
    Institution +
    Degreelevel +
    Cohortyear + 
    Graduates + 
    Twentyfifthpercentile + 
    Seventyfifthpercentile + 
    Percentfulltime + 
    Majorgroup,  data = df_nn)
m <- clean_names(m)

#Excluding Major and Major code made the NN run way smoother. The inclusion of Major caused too many dummy variables to be created and using Major Group in its place is better for generalizing. 


#Normalize the data 
maxs <- apply(m, 2, max)  
mins <- apply(m, 2, min) 
scaled <- as.data.frame(scale(m, center = mins, scale = maxs - mins)) 
  
#Split the data into training and testing set 
index <- sample(1:nrow(m), round(0.75 * nrow(m))) 
train_ <- scaled[index,] 
test_ <- scaled[-index,]


#Create Formula
col_list <- paste(c(colnames(train_[,-c(1,53)])),collapse="+")
col_list <- paste(c("median_earnings~",col_list),collapse="")
f <- formula(col_list)

#Build Neural Network
nn <- neuralnet(f, data = train_, 
                algorithm = "rprop+",
                hidden=c(22, 12),
                threshold=0.5,
                rep=5,
                lifesign = "minimal",
                linear.output = TRUE)
  

#Predict on test data 
pr.nn <- compute(nn, test_[,1:52]) 

#Compute mean squared error 
pr.nn_ <- pr.nn$net.result * (max(scaled$"median_earnings") - min(scaled$"median_earnings")) + min(scaled$"median_earnings")

test.r <- (test_$median_earnings) * (max(scaled$median_earnings) - min(scaled$median_earnings)) + min(scaled$median_earnings) 

MSE.nn <- sum((test.r - pr.nn_)^2) / nrow(test_) 

#Plot the neural network 
plot(nn)

#Regression line
plot(test_$median_earnings, pr.nn_, col = "red",  main = 'Real vs Predicted') 
abline(0, 1, lwd = 2)

#nn$result.matrix
```

```{r}
merged[merged == "**"] <- NA
merged$"Major group"[is.na(merged$"Major group")] <- 'Library Science'
merged$STEM_or_Non_STEM <- as.factor(ifelse(merged$"Major group" == "Engineering", 'STEM',
                     ifelse(merged$"Major group" == "Computer & Information Sciences", 'STEM',
                     ifelse(merged$"Major group" == "Engineering Technician", 'STEM',
                     ifelse(merged$"Major group"== "Mathematics & Statistics", 'STEM',
                     ifelse(merged$"Major group"== "Biological Sciences", 'STEM',
                     ifelse(merged$"Major group"== "Natural Resources", 'STEM','Other')))))))
merged$"Median earnings: full-time workers" <- as.numeric(merged$"Median earnings: full-time workers")
merged$Median_Earnings <- merged$"Median earnings: full-time workers"

boxplot(Median_Earnings ~ STEM_or_Non_STEM, data = merged)

cohort_order <- c("1 year after graduation", "5 years after graduation", "10 years after graduation")
    merged$"Cohort Year" <- factor(merged$"Cohort Year", levels=cohort_order)

p2 <- ggplot(merged, aes(x=STEM_or_Non_STEM, y=Median_Earnings)) + 
    geom_boxplot() +
    facet_wrap(~merged$`Cohort Year`) + scale_y_continuous(labels = scales::comma)
p2

c1merged <-  filter(merged, merged$"Cohort Year" == "1 year after graduation")
c5merged <-  filter(merged, merged$"Cohort Year" == "5 years after graduation")
c10merged <-  filter(merged, merged$"Cohort Year" == "10 years after graduation")

t.test(data = c1merged, Median_Earnings ~ STEM_or_Non_STEM, mu = 0, alt = "two.sided", conf=0.95, var.eq = F, paired = F)
t.test(data = c5merged, Median_Earnings ~ STEM_or_Non_STEM, mu = 0, alt = "two.sided", conf=0.95, var.eq = F, paired = F)
t.test(data = c10merged, Median_Earnings ~ STEM_or_Non_STEM, mu = 0, alt = "two.sided", conf=0.95, var.eq = F, paired = F)
```